---
title: "Chloride Analysis for Charlie's Paper"
author: "Alene Onion, Matt Kraft"
date: "3/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  {.tabset}

## Map

Sites included in this analysis.

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

# setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/Current")
# source("new_database/Reading.LMAS.Data.R")
# setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/Probability.Sampling")

rm(list=setdiff(ls(), c("newdata")))
library(tidyverse)
draft<-newdata %>% 
  filter(CHARACTERISTIC_NAME=="CHLORIDE") %>% 
  mutate(year=as.numeric(substr(SAMPLE_DATE,1,4))) %>% 
  filter(year>2001) %>% 
  filter(RSLT_VALIDATOR_QUALIFIER!="R"|is.na(RSLT_VALIDATOR_QUALIFIER)) %>% 
  mutate(RSLT_RESULT_VALUE=ifelse(RSLT_VALIDATOR_QUALIFIER=="U",RSLT_METHOD_DETECTION_LIMIT/2,RSLT_RESULT_VALUE))%>%  
  filter(!is.na(RSLT_RESULT_VALUE)) %>% 
  select(LAKE_HISTORY_ID,LOCATION_HISTORY_ID,LOCATION_WATERBODY_CLASSIFICATION,LOCATION_X_COORDINATE,LOCATION_Y_COORDINATE,SAMPLE_DATE,RSLT_RESULT_VALUE) %>% distinct()

library(ggmap)
library(ggrepel)
library(rgdal)
library(sp)
library(spatialEco)
#pulling Chloride ecoregions

sites<-draft %>% select(LAKE_HISTORY_ID,LOCATION_HISTORY_ID,LOCATION_X_COORDINATE,LOCATION_Y_COORDINATE) %>% distinct()
chl<-readOGR("C:/Users/leneo/Dropbox/Alene/Rscripts/SALT_Task_Force/ny_eco_l3","ny_eco_l3")

#change coords to web mercator for the map
#all of our layers are NAD83 so have to convert to web mercator
chl<-sp::spTransform(chl, sp::CRS("+proj=longlat +datum=WGS84 +no_defs"))

  #convert sites to a points layer and then do left join with chls layer
coordinates(sites)=~LOCATION_X_COORDINATE+LOCATION_Y_COORDINATE
proj4string(sites)<-CRS("+proj=longlat")
#first, turn off s2 processing
sf::sf_use_s2(FALSE)
chl_merge<-point.in.poly(sites,chl)

#convert attribute table to a data frame
chl_data<-fortify(chl_merge@data)
chl_data<-chl_data %>% select(LAKE_HISTORY_ID,LOCATION_HISTORY_ID,agg_L3_Key) %>% distinct()
#generate sites list again
draft<-merge(draft,chl_data,by=c('LAKE_HISTORY_ID','LOCATION_HISTORY_ID'),all=TRUE)


#now remove those that are outside the map region
draft<-draft %>% 
  filter(!is.na(agg_L3_Key)) %>% 
    mutate(agg_L3_Name=ifelse(agg_L3_Key=="60","8.0a Eastern Temperate Forests (8.1.3)",""),
         agg_L3_Name=ifelse(agg_L3_Key=="83","8.0b Eastern Temperate Forests (8.1.10, 8.1.1)",agg_L3_Name),
         agg_L3_Name=ifelse(agg_L3_Key=="58","5.3 Atlantic Highlands (5.3.1, 5.3.3)",agg_L3_Name),
         agg_L3_Name=ifelse(agg_L3_Key=="59","8.0c Eastern Temperate Forests (8.1.7, 8.3.1, 8.4.1, 8.5.4)",agg_L3_Name),
         agg_state="NYS Combined")
#save a copy for maximum calculations
full<-draft
#and generate median for every sites
draft<-draft %>% 
  group_by(LAKE_HISTORY_ID,LOCATION_HISTORY_ID,agg_L3_Key,agg_L3_Name,agg_state,LOCATION_X_COORDINATE,LOCATION_Y_COORDINATE,LOCATION_WATERBODY_CLASSIFICATION) %>% 
  summarize(medians=median(RSLT_RESULT_VALUE)) %>% 
  ungroup() %>% 
  mutate(medians=as.numeric(medians))
#pull maximum value for every site
full<- full %>% 
  group_by(LAKE_HISTORY_ID,LOCATION_HISTORY_ID,agg_L3_Key,agg_L3_Name,LOCATION_X_COORDINATE,LOCATION_Y_COORDINATE,LOCATION_WATERBODY_CLASSIFICATION) %>% 
  summarize(maxvalue=max(RSLT_RESULT_VALUE)) %>% 
  ungroup() %>% 
  mutate(maxvalue=as.numeric(maxvalue),
         agg_state="NYS Combined")


rm(list=setdiff(ls(), c("newdata",'draft','full')))
```



```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
library(ggmap)
library(ggrepel)

sites_map<-draft %>% select(LAKE_HISTORY_ID,LOCATION_HISTORY_ID,LOCATION_X_COORDINATE,LOCATION_Y_COORDINATE,agg_L3_Key) %>% distinct()

nybox<-make_bbox(sites_map,lon=LOCATION_X_COORDINATE,lat=LOCATION_Y_COORDINATE)

print(ny.map1<-qmap(nybox,source="osm",maptype="terrain",color="bw")+
  geom_point(data=sites_map,aes(x=LOCATION_X_COORDINATE,y=LOCATION_Y_COORDINATE,label=LAKE_HISTORY_ID,colour=agg_L3_Key),size=4))


DT::datatable(sites_map, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
rm(list=setdiff(ls(), c("newdata",'draft','full')))
```

## Violin Plot

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
 library(ggplot2)
# ggplot(draft,aes(x=agg_L3_Name,y=medians)) +
#   geom_violin()+
#   scale_y_continuous(trans='log10')

#create these files for Charlie's plot
cl_historic_ADJ.agg.summary<-draft %>% 
  group_by(agg_L3_Name) %>% 
  summarize(MEDIAN_COUNT=n()) %>% 
  ungroup() %>% 
  rename(Aggregate_Level = agg_L3_Name) 
cl_historic_ADJ.agg.state<-cl_historic_ADJ.agg.summary %>% 
  mutate(agg_state="NYS Combined") %>% 
  group_by(agg_state) %>% 
  summarize(MEDIAN_COUNT=sum(MEDIAN_COUNT)) %>% 
  ungroup()

#load libraries:
library(devtools)
library(ggplot2)
library(cowplot)
library(pgirmess)
library(grid)


## Create plot of state
#----
## Create boxplot for entire state
plot_eco_state <- ggplot(draft,
aes(x = agg_state,y = as.numeric(medians))) +
geom_violin(width = 1,
color = "grey40") +
stat_boxplot(color = "black",
geom = 'errorbar',
width = 0.075,
size = 1.2) +
geom_boxplot(color = "black",
fill = "#ffffff",
outlier.size = 3,
outlier.shape = 16,
outlier.color = "black",
notch = F,
width = 0.15) +
geom_text(data=cl_historic_ADJ.agg.state,
aes(x=agg_state,
y=0.05,
label=paste("n =",
MEDIAN_COUNT)),
size = 4) +
#xlab("") +
ylab("Cl Concentration (mg/L)") +
geom_hline(yintercept = 250, linetype='dashed', col = '#333333', size = 1.25) +
scale_y_log10(position = "right",
breaks = sort(c(0, 1.0, 10.0, 100.0, 250.0, 1000.0))) +
scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 5)) +
#theme_grey() +
theme(text=element_text(size=15, family="sans", color = "black"),
axis.text = element_text(color = "black"),
axis.title = element_text(face = "bold"),
axis.title.x = element_blank(),
panel.background = element_rect(fill = "darkgrey"),
panel.border = element_rect(color = "black", fill = NA, size = 1.5),
panel.grid.major.y = element_line(color = "white", size = 0.5),
panel.grid.minor.y = element_blank(),
panel.grid.minor.x = element_blank(),
panel.grid.major.x = element_blank(),
plot.title = element_text(hjust = 0.5))



#----
## Create plot of eco regions
#----
## Sort the df to plot concentrations lowest to highest (left to right)
sort_eco_cl_h <- draft %>%
select(agg_L3_Name,
medians) %>%
rename(Aggregate_Level = agg_L3_Name)



sort_eco_cl_h$Aggregate_Level <- as.factor(sort_eco_cl_h$Aggregate_Level)



sort_eco_cl_h <- sort_eco_cl_h %>%
mutate(Aggregate_Level = fct_reorder(Aggregate_Level,
as.numeric(medians)))



plot_eco_sep <- ggplot(sort_eco_cl_h,
aes(x = Aggregate_Level,
y = as.numeric(medians))) +
geom_violin(width = 1.1,
color = "grey40") +
stat_boxplot(color = "black",
geom = 'errorbar',
width = 0.075,
size = 1.2) +
geom_boxplot(color = "black",
fill = "#ffffff",
outlier.size = 3,
outlier.shape = 16,
outlier.color = "black",
notch = F,
width = 0.1) +
geom_text(data=cl_historic_ADJ.agg.summary,
aes(x=factor(Aggregate_Level),
y= 0.05,
label=paste("n =",
MEDIAN_COUNT)),
size = 4) +
#xlab("EcoRegion") +
ylab("Cl Concentration (mg/L)") +
geom_hline(yintercept = 250, linetype='dashed', col = '#333333', size = 1.25) +
scale_y_log10(position = "left",
breaks = sort(c(0, 1.0, 10.0, 100.0, 250.0, 1000.0))) +
scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15)) +
theme_grey() +
theme(text=element_text(size=15, family="sans", color = "black"),
axis.text = element_text(color = "black"),
axis.title = element_text(face = "bold"),
axis.title.x = element_blank(),
panel.background = element_rect(fill = "darkgrey"),
panel.border = element_rect(color = "black", fill = NA, size = 1.5),
panel.grid.major.y = element_line(color = "white", size = 0.5),
panel.grid.minor.y = element_blank(),
panel.grid.minor.x = element_blank(),
panel.grid.major.x = element_blank(),
plot.title = element_text(hjust = 0.5))



plot_row <- plot_grid(plot_eco_sep, plot_eco_state,
rel_widths = c(0.80, 0.20),
align = "h")
library(grid)
library(gridExtra)
x.grob <- grid::textGrob("NYS Cl Ecoregion",
gp=gpar(fontface="bold", family="sans", col="black", fontsize=15))



#add to plot



plot_row <- grid.arrange(arrangeGrob(plot_row, bottom = x.grob))
plot_final <- arrangeGrob(arrangeGrob(plot_row, bottom = x.grob))



# now add the title
# boxplot_title <- ggdraw() +
# draw_label("Median Cl Concentrations (mg/L) for Aggregated EPA Level III EcoRegions",
# fontface = 'bold',
# x = 0.025,
# y = 1,
# hjust = 0
# ) +
# draw_label("and NY State (2001 - 2019)",
# fontface = 'bold',
# x = 0.025,
# y = 0.5,
# hjust = 0) +
# theme(
# # add margin on the left of the drawing canvas,
# # so title is aligned with left edge of first plot
# plot.margin = margin(t = 5, r = 0, b = 0, l = 0,
# unit = "mm"),
# text=element_text(size=14, family="sans")
# )



# plot_grid(
# boxplot_title, plot_row,
# ncol = 1,
# # rel_heights values control vertical title margins
# rel_heights = c(0.1, 1)
# )



# ggsave(filename = "20210723_NY_state_and_Eco_Region_Cl_violin_boxplot_separated_log_grey_fill_png_v1.png", plot_row,
# path = file.path("D:",
# "cstoll_root",
# "Chloride_Project",
# "2020",
# "Figures",
# "boxplot"),
# scale = 1,
# width = 11,
# height = 8.5,
# units = "in",
# dpi = 600)

print(plot_row)  
rm(list=setdiff(ls(), c("newdata",'draft','full')))

```

## tables

General Statistics
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
stats<-draft %>% 
  group_by(agg_L3_Name) %>% 
  mutate(n_Sites=n()) %>% 
  ungroup() %>% 
  group_by(agg_L3_Name,n_Sites) %>% 
  summarise_at(vars(medians),
               list(Median=median, Maximum=max,
                    `Upper Quantile`=~quantile(., probs = 0.75),
                    `Lower Quantile`=~quantile(., probs = 0.25))) %>% 
  ungroup() %>% 
  rename(`Aggregate Level`=agg_L3_Name)
stats1<-draft %>% 
  group_by(agg_state) %>% 
  mutate(n_Sites=n()) %>% 
  ungroup() %>% 
  group_by(agg_state,n_Sites) %>% 
  summarise_at(vars(medians),
               list(Median=median, Maximum=max,
                    `Upper Quantile`=~quantile(., probs = 0.75),
                    `Lower Quantile`=~quantile(., probs = 0.25))) %>% 
  ungroup() %>% 
  rename(`Aggregate Level`=agg_state)
stats<-merge(stats,stats1,all=TRUE)
DT::datatable(stats, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
rm(list=setdiff(ls(), c("newdata",'draft','full')))

```

Percentage of sites with at least one excursion.
This is simply the number of sites with excursions divided by the total number of sites and multiplied by 100.
For class A waters, the excursion level is 250 mg/L.
When I examined all class waters, I used EPA's 304a acute threshold at 230 mg/L.
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
excursion<-full %>% 
  mutate(LOCATION_WATERBODY_CLASSIFICATION=
           ifelse(grepl("A",LOCATION_WATERBODY_CLASSIFICATION),"A","non-A"),
         excursion=ifelse(maxvalue>250,1,0)) %>% 
  rename(class=LOCATION_WATERBODY_CLASSIFICATION,
         `Aggregate Level`=agg_L3_Name) %>% 
  select(`Aggregate Level`,LAKE_HISTORY_ID,class,excursion) %>% distinct() %>% 
  group_by(`Aggregate Level`,class) %>% 
  summarize(frequency=signif((mean(excursion)*100),3)) %>% 
  ungroup() %>% 
  spread(class,frequency)
excursion1<-full %>% 
  mutate(LOCATION_WATERBODY_CLASSIFICATION=
           ifelse(grepl("A",LOCATION_WATERBODY_CLASSIFICATION),"A","non-A"),
         excursion=ifelse(maxvalue>250,1,0)) %>% 
  rename(class=LOCATION_WATERBODY_CLASSIFICATION,
         `Aggregate Level`=agg_state) %>% 
  select(`Aggregate Level`,LAKE_HISTORY_ID,class,excursion) %>% distinct() %>% 
  group_by(`Aggregate Level`,class) %>% 
  summarize(frequency=signif((mean(excursion)*100),3)) %>% 
  ungroup() %>% 
  spread(class,frequency)
excursion<-merge(excursion,excursion1,all=TRUE)
excursion_all<-full %>% 
  mutate(excursion=ifelse(maxvalue>250,1,0)) %>% 
  rename(`Aggregate Level`=agg_L3_Name) %>% 
  select(`Aggregate Level`,LAKE_HISTORY_ID,excursion) %>% distinct() %>% 
  group_by(`Aggregate Level`) %>% 
  summarize(frequency=signif((mean(excursion)*100),3)) %>% 
  ungroup() %>% 
  mutate(class="All Class Waters") %>% 
  spread(class,frequency)
excursion_all1<-full %>% 
  mutate(excursion=ifelse(maxvalue>230,1,0)) %>% 
  rename(`Aggregate Level`=agg_state) %>% 
  select(`Aggregate Level`,LAKE_HISTORY_ID,excursion) %>% distinct() %>% 
  group_by(`Aggregate Level`) %>% 
  summarize(frequency=signif((mean(excursion)*100),3)) %>% 
  ungroup() %>% 
  mutate(class="All Class Waters") %>% 
  spread(class,frequency)
excursion_all<-merge(excursion_all,excursion_all1,all=TRUE)
excursion<-merge(excursion,excursion_all,all=TRUE)

DT::datatable(excursion, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
rm(list=setdiff(ls(), c("newdata",'draft','full')))
```

## Sampling Frequencies

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
	junk<-newdata %>% 
	  filter(!is.na(LAKE_HISTORY_ID),
	         SAMPLE_DATE>'2001-01-01',
	         CHARACTERISTIC_NAME=="CHLORIDE") %>% distinct()
	#removing those that are rejected
	junk5<-newdata %>% 
	  filter(!is.na(LAKE_HISTORY_ID),
	         SAMPLE_DATE>'2001-01-01',
	         CHARACTERISTIC_NAME=="CHLORIDE")%>% distinct() %>% 
	  filter(is.na(RSLT_VALIDATOR_QUALIFIER)|RSLT_VALIDATOR_QUALIFIER!="R")
	#unique locations
	junk6<-junk5 %>% select(LOCATION_HISTORY_ID) %>% distinct()
	#frequency of non-detect
junk7<-junk5 %>% filter(grepl("U",RSLT_LABORATORY_QUALIFIER))
#number of sites with multiple observations
junk8<-junk5 %>% 
  group_by(LAKE_HISTORY_ID) %>% 
  summarize(n=n()) %>% 
  ungroup() %>% 
  filter(n>1) %>% distinct()
```

Number of chloride samples
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
print(nrow(junk))
```

number of chloride samples that don't have rejected values
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
print(nrow(junk5))
```

number of locations with accepted data
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
print(nrow(junk6))
```

number of accepted samples that were non-detects
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
print(nrow(junk7))
```

Number of locations with multiple samples
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
print(nrow(junk8))
```